Ancestry linkage in hybrid zones

### 2016-01-28 drafts/hybrid_zone.tex
a bunch of new figures

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}. When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting in a hybrid zone \citep{Barton1985}. If genetic material can be exchanged via hybridization, clines in ancestry throughout the genome may diminish over time; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which diminishes with physical genomic distance \citep{Barton1986}.\\

The mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}. Here we investigate expected genome-wide patterns of co-ancestry in the presence of selection. Selection against migrant ancestries is expected to reduce admixture proportions at and surrounding a target of selection. The effect on patterns of ancestral block lengths is, however, less well understood. Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.\\

@@

Many population genetic models of hybrid zones have focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

 Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci. It is our goal here to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations. This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state. \\

@@

\subsection{Data}
We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m. musculus/M. m. domesticus} \cite{Turner2011,Turner2014}. The genomic data generated by \citep{Turner2014} was for a mapping population generated by mating wild-caught individuals. Because some individuals were mated multiple times, the mapping population comprises many closely related individuals. We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals). Approximately 280,000 SNPs were represented in the final dataset.

Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\

Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization. Specifically, the high LD between quite distantly linked SNPs in the FS population suggests that a relatively high proportion of early-generation hybrids, whereas the rapid decay of LD in the the TU population reflects a much older hybridization event. Importance of geographic distribution of popgen patterns.

\begin{table}
\begin{tabular}{l|c|c}
Individual ID  & Population \citep{Turner2011} & Longitude(E)\\
\citep{Turner2014}& \citep{Turner2011}&\\
\hline
FP112 & FS&11.665\\
FP67 & FS &11.665\\
FP114 & FS &11.665\\
FP79 & FS &11.665\\
FP152 &  FS &11.665\\
FP11 & GL &11.965\\
FP111&GL &11.965\\
FP59 & HA  &11.721\\
FP92 & HA &11.721\\
FP7 &  HA  & 11.721\\
FP45 & HA&11.721\\
FP244& HO &11.693\\
FP9&   RE&11.994\\
FP5 &  RF&11.767\\
FP78 &  SO &11.539\\
FP52 &  SO &11.539\\
FP180& SO &11.539\\
FP14 & ST&11.539\\
FP40 & ST&11.539\\
FP145 & TU&11.748\\
FP29&  TU&11.748\\
FP135& TU&11.748\\

\end{tabular}
\end{table}

>>>>

\section{Introduction}
The process of speciation commonly involves allopatric divergence in relative isolation \cite{Coyne2004}. When formerly isolated populations come back into contact, a gradient of ancestry can form in geographic space, resulting a hybrid zone \citep{Barton1985}. In addition to steep clines in ancestry, recently formed hybrid zones are characterized by a large physical scale of linkage disequilibrium in ancestry (ancestry-LD). If genetic variation can be exchanged between species via hybridization, both of these signals diminish over time throughout the genome; however, subsets of the genome that are under selection may be expected to maintain stable clines in ancestry as selection against migrant, or hybrid, genotypes impedes gene flow across the hybrid zone \citep{Barton1979a}. The influence of selection on linked loci also prevents the mixing of ancestry at regions surrounding these loci, the effect of which decreases with physical genomic distance \citep{Barton1986}.\\

Within a genome, the mixing of ancestries is brought about by the the process of recombination that takes place every generation. The expected distribution of lengths of unbroken ancestry have been previously described under neutral conditions \cite[e.g.][]{Gravel2012,Sedghifar2015}. Here we investigate expected genome-wide patterns of co-ancestry across a hybrid zone in the presence of selection. Selection against migrant ancestries in a hybrid zone results in reduced frequencies of foreign ancestry surrounding targets of selection. The effect on patterns of ancestral block lengths is, however, less well understood. Because of rapid removal from the population, we expect to observe few short blocks of foreign ancestry surrounding the selected locus, with this effect becoming more pronounced further away from the center of a hybrid zone.\\

@@

While population genetic models of hybrid zones have been extensively explored, these have typically focused on the equilibrium states of ancestry clines \cite{Barton1979a,Barton1986}. These models predict cline width to be narrowest at and nearby selected loci, with cline-width at unliked loci to be wider and more heavily influenced by hybrid zone age and rates of gene flow. As denser genotype data becomes more readily available, genome-wide patterns of linkage can be utilized in the analysis of gene hybrid zones.

 Genome-wide covariance in ancestry can provide information about the number of surviving recombination events that have taken place between ancestral genotypes and therefore provides information about the age of the hybrid zone, and the strength of selection at linked loci. Our goal here is  to understand how strength of selection, density of selected sites, recombination rate and hybrid zone age interact to produce observed patterns of ancestry in present day populations. This may be especially important for an understanding and interpretation of of non-equilibrium patterns of diversity in hybrid zones, as well as understanding the consequences of selection on the extent of mixing of ancestral genomes and, consequently, the maintenance of species distinctions. In the extreme case of zero fitness in first-generation hybrids, for example, paternal genomes will never mix and the stable hybrid zone will resemble a recently formed one. If hybrids have non-zero fitness, and some amount of introgression can occur, and patterns of introgression can provide information about the strength of selection, extent of hybridization, and timing of secondary contact. In this latter case, it may be of additional interest to establish whether the hybrid zone is in its equilibrium state. \\

@@

%\subsection{Data}
%We compare the predictions of our model to genomic data obtained from individuals sampled a naturally occurring hybrid zone between \emph{Mus m. musculus/M. m. domesticus} \cite{Turner2011,Turner2014}. The genomic data generated by \citep{Turner2014} was for a mapping population generated by mating wild-caught individuals. Because some individuals were mated multiple times, the mapping population comprises many closely related individuals. We approximate wild population samples by restricting our analysis to genotyped individuals that had both parents sampled in the same location in such a way that none of the individuals shared a parent (see table for list of individuals). Approximately 280,000 SNPs were represented in the final dataset.

%Ancestry blocks were estimated using ChromopainterV2 \cite{Lawson2012} and STRUCTUREv2 \cite{Falush2003} with $k=2$.\\

%Although we expect the age of this hybrid zone to correspond to historical human movement, and therefore be thousands of years old \cite{Teschke2008}, weighted the physical scale of ancestry-disequilibrium (both in the form of weighted ancestry-LD \citep{Loh2013} and ancestral block lengths) suggest relatively recent hybridization. Specifically, the high LD between quite distantly linked SNPs in the FS population suggests that a relatively high proportion of early-generation hybrids, whereas the rapid decay of LD in the the TU population reflects a much older hybridization event. Importance of geographic distribution of popgen patterns.

%\begin{table}
%\begin{tabular}{l|c|c}
%Individual ID  & Population \citep{Turner2011} & Longitude(E)\\
%\citep{Turner2014}& \citep{Turner2011}&\\
%\hline
%FP112 & FS&11.665\\
%FP67 & FS &11.665\\
%FP114 & FS &11.665\\
%FP79 & FS &11.665\\
%FP152 &  FS &11.665\\
%FP11 & GL &11.965\\
%FP111&GL &11.965\\
%FP59 & HA  &11.721\\
%FP92 & HA &11.721\\
%FP7 &  HA  & 11.721\\
%FP45 & HA&11.721\\
%FP244& HO &11.693\\
%FP9&   RE&11.994\\
%FP5 &  RF&11.767\\
%FP78 &  SO &11.539\\
%FP52 &  SO &11.539\\
%FP180& SO &11.539\\
%FP14 & ST&11.539\\
%FP40 & ST&11.539\\
%FP145 & TU&11.748\\
%FP29&  TU&11.748\\
%FP135& TU&11.748\\

%\end{tabular}
%\end{table}
